---
title: "Hasib"
output: html_document
date: "2026-03-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

Clear enviroment
```{r, include=FALSE}

rm(list = ls(all.names = TRUE)) #clean environment
```

install packages
```{r}
install.packages(c(
  "plyr",
  "tidyverse",
  "rstatix",
  "cocor",
  "RColorBrewer",
  "psych",
  "magrittr",
  "data.table",
  "flextable",
  "officer",
  "flextable",
  "officer",
  "ggplot2"
))
```

upload libraries

```{r, include=FALSE}
library(plyr)
library(tidyverse)
library(rstatix)
library(cocor)
library(RColorBrewer)
library(psych)
library(magrittr )
library(data.table)
library(flextable)
library(officer)
```

```{r}

```




Read files
```{r}

path  <- "~/labrotation.si@gmail.com - Google Drive/My Drive/Hasib Thesis/ExperimentDataConvert/dataAnalysis/dataFiles"

files <-list.files(path, pattern = "\\.csv$", full.names = TRUE, recursive = FALSE)
names(files) <- basename(files)

read_one <- function(f){
  df <- read_csv(f, show_col_types = FALSE,
                 na = c("", "NA", "NaN"),
                 locale = locale(encoding = "UTF-8"),
                 guess_max = 100000)
  if ("ID" %in% names(df)) mutate(df, ID = as.character(.data$ID)) else df
}

DT <- map_dfr(files, read_one, .id = "source_file") %>%
  relocate(source_file, .before = 1)
#QC no of trails
DT%>%
  group_by(ID) %>%
  summarise(n_rows = n(), .groups = "drop")
```

Calculate percentage of each signal_detection category per ID
```{r}

result <- DT %>%
  group_by(ID, signal_detection) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(ID) %>%
  mutate(percentage = 100 * n / sum(n)) %>%
  ungroup()

result

# Sum percentages per ID to verify they equal 100
check_totals <- result %>%
  group_by(ID) %>%
  summarise(total_percentage = sum(percentage), .groups = "drop")

check_totals

# Using the existing 'result' object (with percentages per signal category)

accuracy_from_result <- result %>%
  mutate(response_type = case_when(
    signal_detection %in% c("Hit", "CR") ~ "Correct",
    signal_detection %in% c("FA", "Miss") ~ "Incorrect"
  )) %>%
  group_by(ID, response_type) %>%
  summarise(percentage = sum(percentage), .groups = "drop")

accuracy_from_result
```

Calculate percentage of Hit and Miss for each gapOf 1-11
```{r}

hit_miss_by_gap <- DT %>%
  filter(signal_detection %in% c("Hit", "Miss")) %>%
  group_by(gapOf, signal_detection) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(gapOf) %>%
  mutate(percentage = 100 * n / sum(n)) %>%
  ungroup()

hit_miss_by_gap

# Pivot for easier viewing
hit_pivot <- hit_miss_by_gap %>%
  pivot_wider(names_from = signal_detection, values_from = percentage, id_cols = gapOf)

hit_pivot
```

Calculate percentage of Hit and Miss for each gapOf by condition
```{r}

hit_miss_by_gap_condition <- DT %>%
  filter(signal_detection %in% c("Hit", "Miss")) %>%
  group_by(condition, gapOf, signal_detection) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(condition, gapOf) %>%
  mutate(percentage = 100 * n / sum(n)) %>%
  ungroup()

hit_miss_by_gap_condition

# Pivot for easier viewing
hit_miss_pivot <- hit_miss_by_gap_condition %>%
  pivot_wider(names_from = signal_detection, values_from = percentage, id_cols = c(condition, gapOf))

hit_miss_pivot
```

Calculate percentage of Hit and Miss for each condition
```{r}

hit_miss_by_condition <- DT %>%
  filter(signal_detection %in% c("Hit", "Miss")) %>%
  group_by(condition, signal_detection) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(condition) %>%
  mutate(percentage = 100 * n / sum(n)) %>%
  ungroup()

hit_miss_by_condition

# Pivot for easier viewing
hit_miss_condition_pivot <- hit_miss_by_condition %>%
  pivot_wider(names_from = signal_detection, values_from = c(n, percentage), id_cols = condition)

hit_miss_condition_pivot
```

# Signal Detection Theory Analysis

## Step 1: Calculate Hit Rate for each difficulty level (gapOf 1-11)

```{r}
# Calculate Hit Rate for each difficulty level
# Hit = "Different" responses to actually different pairs / total different pairs at each level

sdt_results <- data.frame(
  difficulty_level = 1:11,
  total_trials = NA,
  hits = NA,
  hit_rate = NA,
  z_hit = NA
)

for (level in 1:11) {
  different_trials <- DT[DT$gapOf == level, ]
  total_different <- nrow(different_trials)
  hits <- sum(different_trials$decision_choice == "Different")
  Hk <- hits / total_different
  
  # Apply log-linear correction for extreme values
  if (Hk == 1) {
    Hk_corrected <- 1 - 1/(2 * total_different)
  } else if (Hk == 0) {
    Hk_corrected <- 1/(2 * total_different)
  } else {
    Hk_corrected <- Hk
  }
  
  sdt_results$total_trials[level] <- total_different
  sdt_results$hits[level] <- hits
  sdt_results$hit_rate[level] <- Hk
  sdt_results$z_hit[level] <- qnorm(Hk_corrected)
}

sdt_results
```

## Step 2: Calculate False Alarm Rate (gapOf = 0)

```{r}
# FA = "Different" responses to identical pairs / total identical pairs
identical_trials <- DT[DT$gapOf == 0, ]
total_identical <- nrow(identical_trials)
different_responses_to_identical <- sum(identical_trials$decision_choice == "Different")

FA <- different_responses_to_identical / total_identical

# Apply log-linear correction for extreme values
if (FA == 0) {
  z_FA <- qnorm(1 - 1/(2 * total_identical))
} else if (FA == 1) {
  z_FA <- qnorm(1/(2 * total_identical))
} else {
  z_FA <- qnorm(FA)
}

cat("Total identical trials (gapOf = 0):", total_identical, "\n")
cat("'Different' responses to identical:", different_responses_to_identical, "\n")
cat("FA rate:", round(FA, 4), "\n")
cat("z(FA):", round(z_FA, 4), "\n")
```

## Step 3 & 4: Calculate d' for each difficulty level

```{r}
# d' = z(Hit) - z(FA) for each difficulty level
sdt_results$z_fa <- z_FA
sdt_results$d_prime <- sdt_results$z_hit - sdt_results$z_fa

sdt_results
```

## Step 5: Final Results Table

```{r}
# Create final results table
cat("\n=== FINAL SDT RESULTS TABLE ===\n")
cat("Difficulty Level | Total Trials | Hits | Hit Rate | d'\n")
cat("-----------------|--------------|------|----------|--------\n")

for (level in 1:11) {
  cat(sprintf("%16d | %12d | %4d | %8.4f | %6.4f\n",
              level, 
              sdt_results$total_trials[level],
              sdt_results$hits[level],
              sdt_results$hit_rate[level], 
              sdt_results$d_prime[level]))
}
```

## Interpretation

```{r}
# Check 1: Does hit rate increase with difficulty level?
cat("\n=== INTERPRETATION ===\n")
cat("1. Does hit rate increase as the difference between figures increases?\n")

# Calculate correlation between difficulty level and hit rate
correlation <- cor(sdt_results$difficulty_level, sdt_results$hit_rate)
cat("   Correlation (level vs hit rate):", round(correlation, 4), "\n")
cat("   ", ifelse(correlation > 0, "Yes, positive trend observed", "No clear trend"), "\n\n")

# Check 2: Extreme performance (ceiling/floor effects)
cat("2. Extreme performance (ceiling/floor effects):\n")
extreme_levels <- which(sdt_results$hit_rate > 0.9 | sdt_results$hit_rate < 0.1)
if (length(extreme_levels) > 0) {
  cat("   Levels with >90% or <10% hit rate:", paste(extreme_levels, collapse = ", "), "\n")
  cat("   ", paste(sdt_results$hit_rate[extreme_levels], collapse = ", "), "\n\n")
} else {
  cat("   No extreme levels detected.\n\n")
}

# Check 3: Informative middle levels
cat("3. Levels with non-extreme performance (most informative for measuring sensitivity):\n")
informative_levels <- which(sdt_results$hit_rate > 0.1 & sdt_results$hit_rate < 0.9)
cat("   Informative levels:", paste(informative_levels, collapse = ", "), "\n")
cat("   Hit rates:", paste(round(sdt_results$hit_rate[informative_levels], 4), collapse = ", "), "\n")
cat("   d' values:", paste(round(sdt_results$d_prime[informative_levels], 4), collapse = ", "), "\n")
```

# Correlation Plot: gapOf vs d'

```{r}
# Correlation plot: gapOf vs d'

library(ggplot2)

# Create plot
plot_sdt <- ggplot(sdt_results, aes(x = difficulty_level, y = d_prime)) +
  geom_point(size = 3, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(
    title = "Correlation between Difficulty Level (gapOf) and Sensitivity (d')",
    x = "Difficulty Level (gapOf)",
    y = "Sensitivity (d')"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Display plot
plot_sdt

# Calculate correlation
cor_test <- cor.test(sdt_results$difficulty_level, sdt_results$d_prime)
cat("Correlation:", round(cor_test$estimate, 4), "\n")
cat("p-value:", round(cor_test$p.value, 6), "\n")
```

# Correlation Plot: gapOf vs Hit Rate

```{r}
# Correlation plot: gapOf vs Hit Rate
plot_hitrate <- ggplot(sdt_results, aes(x = difficulty_level, y = hit_rate)) +
  geom_point(size = 3, color = "darkgreen") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title = "Correlation between Difficulty Level (gapOf) and Hit Rate",
    x = "Difficulty Level (gapOf)",
    y = "Hit Rate"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Display plot
plot_hitrate

# Correlation test
cor_hitrate <- cor.test(sdt_results$difficulty_level, sdt_results$hit_rate)
cat("Correlation:", round(cor_hitrate$estimate, 4), "\n")
cat("p-value:", round(cor_hitrate$p.value, 6), "\n")
```
# Psychometric Function: Hit Rate vs Difficulty Level

```{r}
# Psychometric Function: Hit Rate vs gapOf
library(ggplot2)

# Fit a logistic model
psychometric_model <- glm(hit_rate ~ difficulty_level, 
                          data = sdt_results, 
                          family = binomial(link = "logit"),
                          weights = total_trials)

# Create prediction data
pred_data <- data.frame(
  difficulty_level = seq(1, 11, by = 0.1)
)
pred_data$predicted <- predict(psychometric_model, 
                                newdata = pred_data, 
                                type = "response")

# Plot with fitted curve
ggplot() +
  # Fitted psychometric curve
  geom_line(data = pred_data, 
            aes(x = difficulty_level, y = predicted), 
            color = "blue", size = 1.5) +
  # Observed data points (sized by number of trials)
  geom_point(data = sdt_results, 
             aes(x = difficulty_level, y = hit_rate, size = total_trials), 
             color = "darkred", alpha = 0.8) +
  scale_y_continuous(limits = c(0, 1), 
                      name = "Hit Rate (P('Different'))") +
  scale_x_continuous(limits = c(0, 12), 
                     name = "Difficulty Level (gapOf)",
                     breaks = 1:11) +
  scale_size_continuous(name = "Number of Trials") +
  labs(title = "Psychometric Function: Hit Rate vs Difficulty Level") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# Model summary
cat("\nLogistic Regression Summary:\n")
summary(psychometric_model)
```

